\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-22 tex/paper.tex
from two to one algorithm

In this note, we develop two likelihood-free inference algorithms
for non-differentiable, implicit generative models. The first allows us
to perform empirical Bayes through variational inference, and the
second provides a point estimator of the parameters $\bftheta$.
We adapt the adversarial
training procedure of generative adversarial
networks~\cite{goodfellow2014generative} by replacing the implicit generative
network with a domain-based scientific simulator, and solve the resulting
non-differentiable minimax problem by minimizing variational upper
bounds~\citep{2011arXiv1106.4487W,2012arXiv1212.4507S} of the adversarial
objectives.
The objective of both algorithms is to match marginal distribution of
the generated data to the empirical distribution of the observations.

@@

Perhaps the closest to our work is~\cite{bernton2017inference}, which uses the
Wasserstein distance both to find point estimates $\bftheta^*$ and as a part of
a rejection sampler in an ABC-like setup (as opposed to variational inference).
They emphasize that this approach allows one to eliminate the summary statistics
typically used in ABC and calculate the Wasserstein distance explicitly, without
making use of the Kantorovich-Rubinstein duality and of a critic $d$. For high-dimensional
data, they note that this is computationally expensive and introduce an
approximation based on projection of the data onto Hilbert space-filling curves.
Their Remark 5.1 points out that~\citep{montavon2016wasserstein} proposed an
approximation of the gradient of an entropy-regularized Wasserstein distance,
which uses a similar duality. They note that ``unfortunately, it is not
applicable in the setting of purely generative models, as it involves point-wise
evaluations of the derivative of the log-likelihood.'' Thus, our contribution is
to provide gradients of an approximate MWE by taking expectations with the
variational program $q(\bfx | \bfpsi)$. This paired with the critic
Kantorovich-Rubinstein dual formulation of the Wasserstein distance
allows us to work in high dimensions without summary statistics and to
optimize $\bfpsi$ with stochastic gradient descent. Additionally,
our AVO-EB procedure avoids the inefficiencies of
their Wasserstein rejection sampler incurred from using the ABC-likelihood.
In the case of small $N$, we note however that fast algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}
are alternatives to adversarial learning that are worth considering.

@@

In this note, we develop likelihood-free inference algorithms for
non-differentiable, implicit generative models. The algorithms combine
adversarial training  with variational optimization to minimize variational
upper bounds  on the otherwise non-differentiable adversarial objectives. The
AVO algorithm enables empirical Bayes through variational inference in the
likelihood-free setting. This approach does not incur the inefficiencies of an
ABC-like rejection sampler. When the model is well-specified, the AVO algorithm 
provides point estimates for
the generative model, which asymptotically corresponds to the data generating
parameters. The algorithm works on continuous or discrete data.

>>>>

In this note, we develop a likelihood-free inference algorithm
for non-differentiable, implicit generative models.
% The first allows us
% to perform empirical Bayes through variational inference, and the
% second provides a point estimator of the parameters $\bftheta$.
We adapt the adversarial
training procedure of generative adversarial
networks~\cite{goodfellow2014generative} by replacing the implicit generative
network with a domain-based scientific simulator, and solve the resulting
non-differentiable minimax problem by minimizing variational upper
bounds~\citep{2011arXiv1106.4487W,2012arXiv1212.4507S} of the adversarial
objectives.
The objective of both algorithms is to match marginal distribution of
the generated data to the empirical distribution of the observations.

@@

Perhaps the closest to our work is~\cite{bernton2017inference}, which uses the
Wasserstein distance both to find point estimates $\bftheta^*$ and as a part of
a rejection sampler in an ABC-like setup (as opposed to variational inference).
They emphasize that this approach allows one to eliminate the summary statistics
typically used in ABC and calculate the Wasserstein distance explicitly, without
making use of the Kantorovich-Rubinstein duality and of a critic $d$. For high-dimensional
data, they note that this is computationally expensive and introduce an
approximation based on projection of the data onto Hilbert space-filling curves.
Their Remark 5.1 points out that~\citep{montavon2016wasserstein} proposed an
approximation of the gradient of an entropy-regularized Wasserstein distance,
which uses a similar duality. They note that ``unfortunately, it is not
applicable in the setting of purely generative models, as it involves point-wise
evaluations of the derivative of the log-likelihood.'' Thus, our contribution is
to provide gradients of an approximate MWE by taking expectations with the
variational program $q(\bfx | \bfpsi)$. This paired with the critic
Kantorovich-Rubinstein dual formulation of the Wasserstein distance
allows us to work in high dimensions without summary statistics and to
optimize $\bfpsi$ with stochastic gradient descent. Additionally,
our procedure avoids the inefficiencies of
their Wasserstein rejection sampler incurred from using the ABC-likelihood.
In the case of small $N$, we note however that fast algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}
are alternatives to adversarial learning that are worth considering.

@@

In this note, we develop a likelihood-free inference algorithm for
non-differentiable, implicit generative models. The algorithm combines
adversarial training  with variational optimization to minimize variational
upper bounds  on the otherwise non-differentiable adversarial objectives. The
AVO algorithm enables empirical Bayes through variational inference in the
likelihood-free setting. This approach does not incur the inefficiencies of an
ABC-like rejection sampler. When the model is well-specified, the AVO algorithm
provides point estimates for
the generative model, which asymptotically corresponds to the data generating
parameters. The algorithm works on continuous or discrete data.
